Methods for Large Scale Learning to Rank

### 2014-02-20 report/BenchmarkResults.tex
hernoeming mappen

The Yahoo! Learning to Rank Challenge consists of two tracks that uses the two datasets respectively: a standard Learning-to-Rank track and a transfer learning track where the goal was to learn a specialized ranking function that can be used for a small country by leveraging a larger training set of another country. For this experiment I will only look at the standard Learning-to-Rank dataset because transfer learning is a separate research area that is not included in this thesis.\\
\begin{table}[!h]
\begin{tabular}{l|lll}
 & Train & Validation & Test \\ 
 \hline
Queries & 19,994 & 2,994 & 6,983 \\ 
Documents & 473,134 & 71,083 & 165,660 \\ 
Features & 519 & - & - \\ 
\end{tabular}
\caption{Yahoo! Learning to Rank Challende dataset characteristics, as described in the challenge overview paper\cite{Chapelle2011a}}
\label{tab:yahoo_characteristics}
\end{table}\\
Both \ac{nDCG} and \ac{ERR} are measured as performance metrics, but the final standings of the challenge were based on the \ac{ERR} values. Model validation on the Learning-to-Rank methods perticipanting in the challenge is performed using a train/validation/test-set split following the characteristics shown in Table \ref{tab:yahoo_characteristics}. Competitors could train on the training set and get immediate feedback on their performance on the validation set. The test set performance is used to create the final standings and is only measured after the competition has ended to avoid overfitting on the test set. The large number of documents, queries and features compared to other benchmark datasets makes the Yahoo! Learning to Rank Challenge dataset interesting. Yahoo did not provide detailed feature descriptions to prevent competitors to get detailed insight in the characteristics of the Yahoo data collection and features used at Yahoo. Instead high level descriptions of feature categories are provided. The following categories of features are described in the challenge overview paper\cite{Chapelle2011a}:\\
\begin{description}
\item[Web graph]{Quality and popularity metrics of web documents, e.g. PageRank\cite{Page1999}}.
\item[Document statistics]{Basic document statistics such as the number of words and url characteristics.}
\item[Document classifier]{Results of various classifiers on the documents. These classifiers amongst others include: spam, adult, language, main topic, and quality classifiers.}
\item[Query]{Basic query statistics, such as the number of terms, query frequency, and click-through rate.}
\item[Text match]{Textual similarity metrics between query and document. Includes \ac{TF-IDF}, BM25\cite{Robertson2009} and other metrics for different sections of the document.}
\item[Topical matching]{These features go beyond similarity at word level and compute similarity on topic level. For example by classifying both the document and the query in a large topical taxonomy.}
\item[Click]{Click-based user feedback.}
\item[External references]{Document meta-information such as Delicious\footnote{https://delicious.com/} tags}
\item[Time]{Document age and historical in- and outlink data that might help for time sensitive queries.}
\end{description}

>>>>

The Yahoo! Learning to Rank Challenge consists of two tracks that uses the two datasets respectively: a standard Learning-to-Rank track and a transfer learning track where the goal was to learn a specialized ranking function that can be used for a small country by leveraging a larger training set of another country. For this experiment I will only look at the standard Learning-to-Rank dataset because transfer learning is a separate research area that is not included in this thesis.\\
\begin{table}[!h]
\begin{tabular}{l|lll}
 & Train & Validation & Test \\ 
 \hline
Queries & 19,994 & 2,994 & 6,983 \\ 
Documents & 473,134 & 71,083 & 165,660 \\ 
Features & 519 & - & - \\ 
\end{tabular}
\caption{Yahoo! Learning to Rank Challenge dataset characteristics, as described in the challenge overview paper\cite{Chapelle2011a}}
\label{tab:yahoo_characteristics}
\end{table}\\
Both \ac{nDCG} and \ac{ERR} are measured as performance metrics, but the final standings of the challenge were based on the \ac{ERR} values. Model validation on the Learning-to-Rank methods participating in the challenge is performed using a train/validation/test-set split following the characteristics shown in Table \ref{tab:yahoo_characteristics}. Competitors could train on the training set and get immediate feedback on their performance on the validation set. The test set performance is used to create the final standings and is only measured after the competition has ended to avoid overfitting on the test set. The large number of documents, queries and features compared to other benchmark datasets makes the Yahoo! Learning to Rank Challenge dataset interesting. Yahoo did not provide detailed feature descriptions to prevent competitors to get detailed insight in the characteristics of the Yahoo data collection and features used at Yahoo. Instead high level descriptions of feature categories are provided. The following categories of features are described in the challenge overview paper\cite{Chapelle2011a}:\\
\begin{description}
\item[Web graph]{Quality and popularity metrics of web documents, e.g. PageRank\cite{Page1999}}.
\item[Document statistics]{Basic document statistics such as the number of words and url characteristics.}
\item[Document classifier]{Results of various classifiers on the documents. These classifiers amongst others include: spam, adult, language, main topic, and quality classifiers.}
\item[Query]{Basic query statistics, such as the number of terms, query frequency, and click-through rate.}
\item[Text match]{Textual similarity metrics between query and document. Includes \ac{TF-IDF}, BM25\cite{Robertson2009} and other metrics for different sections of the document.}
\item[Topical matching]{These features go beyond similarity at word level and compute similarity on topic level. For example by classifying both the document and the query in a large topical taxonomy.}
\item[Click]{Click-based user feedback.}
\item[External references]{Document meta-information such as Delicious\footnote{https://delicious.com/} tags}
\item[Time]{Document age and historical in- and outlink data that might help for time sensitive queries.}
\end{description}

### 2014-02-20 report/Chapters/Background.tex
hernoeming mappen

\section{Pairwise Approach}
\section{Listwise Approach}
\section{Evaluation Methods}
\subsection{Normalized Discounted Cumulative Gain}
\subsubsection{Cumulative Gain}
\subsubsection{Discounted Cumulative Gain}
\ac{DCG}
\subsubsection{Normalized Discounted Cumulative Gain}
\ac{nDCG}
\subsection{Expected Reciprocal Rank}
\ac{ERR}
\subsection{Mean Average Precision}
\ac{MAP}

>>>>

\section{Pairwise Approach}
\section{Listwise Approach}
\chapter{Evaluation Methods}

\section{Normalized Discounted Cumulative Gain}
Cumulative gain, or its predecessor discounted cumulative gain and normalized discounted cumulative gain, is one of the most widely used measures for effectiveness of ranking methods.
\subsection{Discounted Cumulative Gain}
The \ac{DCG}\cite{Burges2005} at a position $p$ is defined as\\
$DCG_p = \sum\nolimits_{i=1}^p \frac{2^{rel_i-1}}{log_2(i+1)}$\\
with $rel_i$ the graded relevance of the result at position $i$. The idea is that highly relevant documents that appear lower in a search result should be penalized (discounted). This discounting is done by reducing the graded relevance  logarithmically proportional to the position of the result.
\subsection{Normalized Discounted Cumulative Gain}
\ac{nDCG} normalizes the \ac{DCG} metric to a value in the [0,1] interval by dividing by the \ac{DCG} value of the optimal rank. This optimal rank is obtained by sorting documents on relevance for a given query. We can write the definition of \ac{nDCG} down mathematically as\\
$nDCG_p = \frac{DCG_p}{IDCG_p}$
\section{Expected Reciprocal Rank}
\ac{ERR}\cite{Chapelle2009} was designed based on the observation that \ac{nDCG} is based on the false assumption that the usefulness of a document at rank $i$ is independent of the usefulness of the documents at rank less than $i$. \ac{ERR} is based on the reasoning that users are likely to stop exploring the result list once they have found a document that satisfied their information need. The \ac{ERR} metric is defined as the expected reciprocal length of time that the user will take to find a relevant document. \ac{ERR} is formally defined as\\
$ERR = \sum\nolimits_{r=1}^n \frac{1}{r} \prod\nolimits_{i=1}^{r-1}(1-R_i)R_i$\\
where the product sequence part of the formula represents the chance that the user will stop at position $r$.\\
An algorithm to compute \ac{ERR} is shown in Listing \ref{lst:err}. The algorithm requires relevance grades $g_i$, $1 \le i \le n$ and mapping function $R$ that maps relevance grades to probability of relevance.
\begin{lstlisting}[caption={Algorithm to compute the ERR metric, obtained from \cite{Chapelle2009}}, label={lst:err},language=Ada]
p <- 1, ERR <- 0
for r=1 to n do
	R <- R(g[r])
	ERR <- ERR + p * R/r
	p <- p * (1-R)
end for
return ERR
\end{lstlisting}
\section{Mean Average Precision}
\ac{AP}\cite{Zhu2004} is an often used binary relevance judgement based metric that can be seen as a trade-off between precision and recall that is defined as\\
$AP = \frac{\sum\nolimits_{k=1}^{n}Precision(k)*relevance(k)}{\text{number of relevant docs}}$\\
With $k$ being the positions in the result set between 1 and $n$.
\ac{MAP} is the average \ac{AP} for a set of queries.\\
$MAP = \frac{\sum\nolimits_{q=1}^{Q}AP(q)}{Q}$

### 2014-02-20 report/RelatedWork.tex
hernoeming mappen

Yan et al\cite{Yan2009,Yan2010,Yan2011,Yan2012} described the development and incremental improvement of a \ac{SIMD} architecture \ac{FPGA} designed to run the Neural Network based LambdaRank Learning-to-Rank algorithm. This architecture achieved a 29.3X speed-up compared to the software implementation when evaluated on data from a commercial search engine. The exploration of \ac{FPGA} for Learning-to-Rank showed other advantages of the \ac{FPGA} approach next to faster model training. In their latest publication\cite{Yan2012} the FPGA based LambdaRank implementation showed it could achieve up to 19.52X power efficiency and 7.17X price efficiency for query processing compared to Intel Xeon servers currently used at the commercial search engine.

@@

De Sousa et al\cite{DeSousa2012} proposed a \ac{GPGPU} approach using the \ac{GPU} both learning the ranking function and for query processing, thereby improving both training time and query time. An association rule based Learning-to-Rank approach proposed by \cite{Veloso2008} has been implemented using the GPU in such a way that the set of rules van be computed in parallel, in different threads, for each document. A speed-up of 127X in query processing time is reported based on evaluation on the LETOR dataset. The speed-up achieved at learning the ranking function was unfortunately not stated.

>>>>

Yan et al\cite{Yan2009,Yan2010,Yan2011,Yan2012} described the development and incremental improvement of a \ac{SIMD} architecture \ac{FPGA} designed to run the Neural Network based LambdaRank Learning-to-Rank algorithm. This architecture achieved a 29.3X speed-up compared to the software implementation when evaluated on data from a commercial search engine. The exploration of \ac{FPGA} for Learning-to-Rank showed other advantages of the \ac{FPGA} approach next to faster model training. In their latest publication\cite{Yan2012} the \ac{FPGA} based LambdaRank implementation showed it could achieve up to 19.52X power efficiency and 7.17X price efficiency for query processing compared to Intel Xeon servers currently used at the commercial search engine.

@@

De Sousa et al\cite{DeSousa2012} proposed a \ac{GPGPU} approach using the \ac{GPU} both learning the ranking function and for query processing, thereby improving both training time and query time. An association rule based Learning-to-Rank approach proposed by \cite{Veloso2008} has been implemented using the \ac{GPU} in such a way that the set of rules van be computed in parallel, in different threads, for each document. A speed-up of 127X in query processing time is reported based on evaluation on the LETOR dataset. The speed-up achieved at learning the ranking function was unfortunately not stated.
